## Example 3

<!--
Set the working directory.
No Rchunk is needed to accomplish this.
-->

<!--
Read in the example3.csv data. For each of the following steps you
should preface your R commands with some plain English explanation.
-->

The first step is to read in the data.
```{r
dat<-read.csv("example3.csv")
```
<!--
Check the structure of the data.
-->

The next step is to check out the structure

```{r}
str(dat)
```

As we can see dat is a data frame with three variables: gender,
political party affiliation, and support for some initiative.

<!--
Find the mean, median, and variance for the overall level of
support.
-->

Now we'll find some basic summary statistics including the
mean, median, and variance.

```{r}
mean(dat$support)
median(dat$support)
var(dat$support)
```

From the previous analyses, we can see that the mean level of
support is 63.4, the median is 67.0, and the variance is 171.6.

<!--
Find the mean level of support for males only.
-->

Now, let's see what the mean level of support is for males.

```{r}
mean(dat$support[dat$gender=="male"])
```

Here we can see that the mean for males is 58.5.

<!--
Find the mean level of support for democratic females only.
-->

Now, let's see what the mean level of support is for democratic
females.

```{r}
mean(dat$support[dat$gender=="female" & dat$party=="democrat"])
```

Here we can see that the mean for democratic females is 72.1.