## Homework 3 

<!--
You can use the example3.Rmd or other files to find examples
of the codes you'll need to use to complete this assignment. I strongly
urge you to copy and paste relevant code that you know works from my
examples or instructions and then edit that code so you're working
with the appropriate variables for the homework.
-->
go into your working directory and select the appropriate file (homework3.Rmd). Then go to session, set working directory, and source file.
```{r}
dat<-read.csv("homework3.Rmd")
```
This enables r studio to read the file
<!--
Read in the homework3.csv data. For each of the following steps you
should preface your R commands with some plain English explanation.
-->
<!--
Check the structure of the data.
-->
```{r}
str(dat)
```
This gives us a description of the data
data.frame':	100 obs. of  3 variables:
 $ major       : Factor w/ 2 levels "business","psychology": 2 2 2 2 2 2 2 2 2 2 ...
 $ year        : Factor w/ 4 levels "freshman","junior",..: 1 1 1 1 1 1 1 1 1 1 ...
 $ conservatism: num  44.5 38.6 34.2 32.2 42.3 ...
> 
<!--
Find the mean, median, and standard deviation using the sd() function
for the overall level of conservatism.
--> 
```{r}
mean(dat$conservatism)
``` 
The mean of overall conservatism is 52.29458
```{r}
sd(dat$conservatism)
``` 
The standard deviation for overall conservatism is 19.75757
```{r}
median(dat$conservatism)
``` 
The median of overall conservatism is 53.83231

<!--
Find the mean level of support for freshmen only.
-->
```{r}
mean(dat$conservatism[dat$year=="freshman"])
```
The mean level of support for freshmen is 47.50033

<!--
Find the mean level of support for male business students only.
-->
```{r}
mean(dat$conservatism[dat$year=="freshman"&dat$major=="business"])
```
The mean level of support for male business students is 61.60146
