## Example 6

<!--
Set the working directory.
No Rchunk is needed to accomplish this.
-->

<!--
Read in the example6.csv data. For each of the following steps you
should preface your R commands with some plain English explanation.
-->

The first step is to read in the data.
```{r}
dat<-read.csv("example6.csv")
```
<!--
Check the structure of the data and change any variables that should
be factor variables into factor variables
-->

The next step is to check out the structure

```{r}
str(dat)
```

From the structure we can see that position, subject, and subject2
should all be factor variables. So let's fix that.

```{r}
dat$position<-as.factor(dat$position)
dat$subject<-as.factor(dat$subject)
dat$subject2<-as.factor(dat$subject2)
```

<!--
Perform a one-way between-subjects ANOVA examining for effects of 
the list position when participants have an intervening distractor
task to perform before recalling the lists they just learned.
Also create a line graph to help you interpret the results of your
analysis.
-->

First we need to get only the data from the conjunction searches.

```{r}
temp<-dat[dat$distractor=="yes",]
```

Now we can run our ANOVA.

```{r}
summary(aov(percent~position, data=temp))
```

There's a significant effect of which position a word appears in on
a list, F(9,110) = 16.3, p < .05. To interpret these data better,
I'll create a line graph that shows the percent recalled for the
various list positions when participants are distracted.

```{r}
library(ggplot2)
library(gplots)
library(dplyr)

temp<-temp%>%group_by(position)%>%
  summarize(means=mean(percent),
            sems=sd(percent)/sqrt(length(percent)))

f<-ggplot(temp, aes(x=as.factor(position),
                    y=means,
                    group=1))+
  geom_line(size=1)+
  geom_point(size=2)+
  geom_errorbar(aes(ymax=means+sems,
                    ymin=means-sems),
                width=.2)+
  ggtitle("Serial Position Effect")+
  labs(x="List Position",y="Percent Recalled")+
  theme(plot.title=element_text(size=15,face="bold",vjust=.5))+
  theme(axis.title.x=element_text(size=12,face="bold",vjust=-.25))+
  theme(axis.title.y=element_text(size=12,face="bold",vjust=1))+
  theme(axis.text.x=element_text(size=10,face="bold",color="black"))+
  theme(axis.text.y=element_text(size=10,face="bold",color="black"))+
  coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),
                         max(temp$means)+2*max(temp$sems)))+
  theme(panel.border=element_blank(),axis.line=element_line())+
  theme(panel.grid.major.x=element_blank())+
  theme(panel.grid.major.y=element_line(color="darkgrey"))+
  theme(panel.grid.minor.y=element_blank())
f
```
From the figure we can see that participants do the best on the
words that occur at the beginning of the list and worst on the words
that occur at the ending of the list.

<!--
Perform a one-way within subjects ANOVA on the same data you analyzed
earlier. Tip: use subject as the subjects variable for this analysis.
-->

Because I changed temp, I don't have the same data I analyzed last
time, so I'm going to reuse the code I first used to create temp
so I can perform the appropriate analysis.

```{r}
temp<-dat[dat$distractor=="yes",]
```

Now I'll perform the within-subjects analysis.

```{r}
summary(aov(percent~position+
              Error(subject/position),data=temp))
```

Participants performed better on earlier list positions than they
did on later list positions, F(9,99) = 51.0, p = .05.

<!--
Perform a between-subjects ANOVA examining for the effects of and
interactions between list position and whether or not the
participants had an intervening distractor task before recalling
the lists of words.
-->

Now, I'll perform a between-subjects 2-way ANOVA looking for effects
of and interactions between list position and whether or not
participants completed an intervening task.

```{r}
summary(aov(percent~position*distractor, data=dat))
```

The analyses show that there was an effect of position, F(9,220) =
22.7, p < .05. There was also an effect of distractor, F(1,220) =
31.3, p < .05. There was also an interaction between position and
distractor, F(9,220) = 8.07, p < .05.

<!--
Create a line graph and a grouped bar graph to help us interpret
the results of the previous analyses. Include error bars in your
figures.
-->

To help me interpret the earlier analyses, I'll create line and
grouped bar graphs for the data I analyzed. But first I need to
summarize the data.

```{r}
temp<-dat%>%group_by(position,distractor)%>%
  summarize(means=mean(percent),
            sems=sd(percent)/sqrt(length(n)))
```

Now I can create the figures.

```{r}
f<-ggplot(temp, aes(x=position,
                    y=means,
                    group=distractor,
                    color=distractor))+
  geom_line(size=1)+
  geom_point(size=2)+
  scale_color_manual(values=c("darkgreen","red"),
                     name="Distractor",
                     breaks=c("no","yes"),
                     labels=c("No", "Yes"))+
  geom_errorbar(aes(ymax=means+sems, ymin=means-sems),width=.2)+
  ggtitle("Serial Position With and Without Distraction")+
  labs(x="List Position",y="Percent Recalled")+
  theme(plot.title=element_text(size=15,face="bold",vjust=.5))+
  theme(axis.title.x=element_text(size=12,face="bold",vjust=-.25))+
  theme(axis.title.y=element_text(size=12,face="bold",vjust=1))+
  theme(axis.text.x=element_text(size=10,face="bold",color="black"))+
  theme(axis.text.y=element_text(size=10,face="bold",color="black"))+
  coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),
                         max(temp$means)+2*max(temp$sems)))+
  theme(panel.border=element_blank(),axis.line=element_line())+
  theme(panel.grid.major.x=element_blank())+
  theme(panel.grid.major.y=element_line(color="darkgrey"))+
  theme(panel.grid.minor.y=element_blank())+
  theme(legend.position=c(.4,.76))+
  theme(legend.background=element_blank())+
  theme(legend.background=element_rect(color="black"))+
  theme(legend.title=element_blank())+
  theme(legend.title=element_text(size=12))+
  theme(legend.title.align=.5)+
  theme(legend.text=element_text(size=10,face="bold"))
f

f<-ggplot(temp, aes(x=position,
                    y=means,
                    fill=distractor))+
  geom_bar(stat="identity",position=position_dodge())+
  scale_fill_manual(values=c("darkgreen","red"),
                     name="Distractor",
                     breaks=c("no","yes"),
                     labels=c("No", "Yes"))+
  geom_errorbar(aes(ymax=means+sems,
                    ymin=means-sems),
                width=.2,
                position=position_dodge(.9))+
  ggtitle("Serial Position With and Without Distraction")+
  labs(x="List Position",y="Percent Recalled")+
  theme(plot.title=element_text(size=15,face="bold",vjust=.5))+
  theme(axis.title.x=element_text(size=12,face="bold",vjust=-.25))+
  theme(axis.title.y=element_text(size=12,face="bold",vjust=1))+
  theme(axis.text.x=element_text(size=10,face="bold",color="black"))+
  theme(axis.text.y=element_text(size=10,face="bold",color="black"))+
  coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),
                         max(temp$means)+2*max(temp$sems)))+
  theme(panel.border=element_blank(),axis.line=element_line())+
  theme(panel.grid.major.x=element_blank())+
  theme(panel.grid.major.y=element_line(color="darkgrey"))+
  theme(panel.grid.minor.y=element_blank())+
  theme(legend.position=c(.4,.76))+
  theme(legend.background=element_blank())+
  theme(legend.background=element_rect(color="black"))+
  theme(legend.title=element_blank())+
  theme(legend.title=element_text(size=12))+
  theme(legend.title.align=.5)+
  theme(legend.text=element_text(size=10,face="bold"))
f
```

<!--
Interpret the statistical results in light of the figures.
-->

From the figures and earlier analyses we can see that people
generally perform best for the beginning and ending words in
a list. We can also see that in general participants performed
worse when they had an intervening distracting task. Most
importantly, the interaction reveals that although individuals
without intervening distracting tasks perform best on the earliest
and latest list positions, those with the distractor tasks
performed better on earlier positions than on later ones.

<!--
Perform a within-subjects two-way ANOVA examining for the effects
of and interactions between list position and distraction. Tip:
use subject as the subject's variable for this analysis. Interpret
the results of the analysis based on the statistical results and
your earlier figures.
-->

Here I'm performing a within-subjects two-way ANOVA to examine for
the effects of and interactions between list position and
distraction.

```{r}
summary(aov(percent~position*distractor+
              Error(subject/(position*distractor)),
                    data=dat))
```

From the analyses and the figures we can see that generally people
perform best for the initial and later list positions, F(9,99) =
71.3, p < .05. We can also see that people are generally more
accurate to recall words when they don't have an intervening
distractor tasks to complete, F(1,11) = 128.4, p < .05. Finally,
we can see that those without a distractor task perform best for the
words at the beginning and end of the lists while, those with a
distractor task perform best for the initial words and worst for the
ending words, F(9,99) = 20.2, p < .05.

<!--
Finally conduct and interpret the results of a mixed ANOVA in which
position varies within subjects and distraction varies between
subjects. Make sure to interpret your results. Tip: use subject2
as your subject's variable for this analysis.
-->

Finally, I'll conduct a mixed ANOVA in which position varies within
participants and distraction varies between participants.

```{r}
summary(aov(percent~position*distractor+
              Error(subject2/position),
            data=dat))
```

From our anlyses, we can see that individuals without distracting
tasks performed better than those who had, F(1,22) = 4.6, p = .0429.
Additionally, participants generally performed better on the words
at the beginning and ends of the lists, F(9,198) = 63.3, p < .05.
Finally, we can see that distracted individuals performed best on
the initial words only while individuals who weren't distracted
performed well on items at the beginning and end of the lists,
F(9,198) = 22.5, p < .05.