## Example 6

<!--
Set the working directory.
No Rchunk is needed to accomplish this.
-->
First we need to set our working directory to homework 6 by selecting the file and selecting set working directory, then to file source location.
<!--
Read in the homework6.csv data. For each of the following steps you
should preface your R commands with some plain English explanation.
The fabricated data are supposed to approximate how people perform
in different types of visual searches. Specifically, we're 
comparing performance in feature searches (e.g. looking for a red
X among green Xs) and conjunction searches (e.g. looking for a red
X among green Xs and red Os). We're also looking at how the number
of distractors within the search display effect our data.
-->
Then, tell r to read the file:
```{r}
dat<-read.csv("homework6.csv")
```


<!--
Check the structure of the data and change any variables that should
be factor variables into factor variables
-->
You will then check the structure of the data.
```{r}
str(dat)
```

We can see that we have 5 variables, 1 numerical, 1 factor, and 3 integers. We will need to change the integers into factor variables with the following command:
```{r}
dat$distractors<-as.factor(dat$distractors)
dat$subject<-as.factor(dat$subject)
dat$subject2<-as.factor(dat$subject2)
```
We need to load our packages with the following commands:
```{r}
library(ggplot2)
library(gplots)
library(dplyr)
```

<!--
Perform a one-way between-subjects ANOVA examining for effects of 
the the number of distractors for the conjunction searches only.
Also create a line graph to help you interpret the results of your
analysis.
-->
Next, we will tell r to only use the data associated with conjunction searches with the following command:
```{r}
temp<-dat[dat$type=="conjunction",]
```
Then we need to summaraze our data
```{r}
temp<-temp%>%group_by(distractors)%>%
  summarize(means=mean(rt),
            sems=sd(rt)/sqrt(length(rt)))
```
Then we set out temp and run our ANOVA
```{r}
temp<-dat[dat$type=="conjunction",]
```
```{r}
summary(aov(rt~distractors, data=temp))
```
In order to make a line graph of our data we will enter the following code:
```{r}
f<-ggplot(temp, aes(x=as.factor(distractors),
                    y=means,
                    group=1))+
  geom_line(size=1)+
  geom_point(size=2)+
  geom_errorbar(aes(ymax=means+sems,
                    ymin=means-sems),
                width=.2)+
  ggtitle("Effect of Distractors in Conjunction Searches")+
  labs(x="Number of Distractors",y="Response Time")+
  theme(plot.title=element_text(size=15,face="bold",vjust=.5))+
  theme(axis.title.x=element_text(size=12,face="bold",vjust=-.25))+
  theme(axis.title.y=element_text(size=12,face="bold",vjust=1))+
  theme(axis.text.x=element_text(size=10,face="bold",color="black"))+
  theme(axis.text.y=element_text(size=10,face="bold",color="black"))+
  coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),
                         max(temp$means)+2*max(temp$sems)))+
  theme(panel.border=element_blank(),axis.line=element_line())+
  theme(panel.grid.major.x=element_blank())+
  theme(panel.grid.major.y=element_line(color="darkgrey"))+
  theme(panel.grid.minor.y=element_blank())
f
```
This graph shows us that as more distractors are presented in conjunction searches, response time increases.

<!--
Perform a one-way within subjects ANOVA on the same data you analyzed
earlier. Tip: use subject as the subjects variable for this analysis.
-->
Now we will perform a one-way between subjects anova examining the effects of the number of distractors for the conjunction searches (finding a red x among green x's and red o's).
```{r}
temp<-dat[dat$type=="conjunction",]
```
```{r}
summary(aov(rt~distractors, data=temp))
```

From our output we can see that there was a significant effect between number of distractors and response time on conjunction searches, as F(2,21) = 8.6, p < .05
<!--
Perform a between-subjects ANOVA examining for the effects of and
interactions between the number of distractors and the search type.
Make sure to interpret your results.
-->
Now I'll perform a between-subjects ANOVA to examine the effects of and interactions between the number of distractors and the search type. First we must set temp to interpret both types of search.
```{r}
temp<-dat[dat$type=="conjunction","feature",]
```
Then we will run a between subjects ANOVA examining for the effects of and interactions between the number of distractors and the search type.
```{r}
summary(aov(rt~distractors*type, data=dat))
```
We can see from this output that there was no significant effect of distractors on rt, as F(2,45) = 2.6, p = .08. There was however a significant effect of search type, F(1,46) = 459.4, p < .05. In addition, there was a main effect between distractors and search type F(2,45) = 7.8, p < .05
<!--
Create a line graph and a grouped bar graph to help us interpret
the results of the previous analyses. Include error bars in your
figures.
-->
In order to create our graphs, we must first summarize the data
```{r}
temp<-dat%>%group_by(distractors,type)%>%
  summarize(means=mean(rt),
            sems=sd(rt)/sqrt(length(rt)))
```
To make the line graph we will use the following code
```{r}
f<-ggplot(temp, aes(x=distractors,
                    y=means,
                    group=type,
                    color=type,))+
  geom_line(size=1)+
  geom_point(size=2)+
  scale_color_manual(values=c("darkgreen","red"),
                     name="Type",
                     breaks=c("conjunction","feature"),
                     labels=c("Conjunction", "Feature"))+
  geom_errorbar(aes(ymax=means+sems, ymin=means-sems),width=.2)+
  ggtitle("Effects of Distractors on Feature and Conjunction Searches")+
  labs(x="Distractors",y="Response Time")+
  theme(plot.title=element_text(size=15,face="bold",vjust=.5))+
  theme(axis.title.x=element_text(size=12,face="bold",vjust=-.25))+
  theme(axis.title.y=element_text(size=12,face="bold",vjust=1))+
  theme(axis.text.x=element_text(size=10,face="bold",color="black"))+
  theme(axis.text.y=element_text(size=10,face="bold",color="black"))+
  coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),
                         max(temp$means)+2*max(temp$sems)))+
  theme(panel.border=element_blank(),axis.line=element_line())+
  theme(panel.grid.major.x=element_blank())+
  theme(panel.grid.major.y=element_line(color="darkgrey"))+
  theme(panel.grid.minor.y=element_blank())+
  theme(legend.position=c(.4,.76))+
  theme(legend.background=element_blank())+
  theme(legend.background=element_rect(color="black"))+
  theme(legend.title=element_blank())+
  theme(legend.title=element_text(size=12))+
  theme(legend.title.align=.5)+
  theme(legend.text=element_text(size=10,face="bold"))
f
```
We can see from the line graph that conjunction search response times were affected more than feature searches in response to number of distractors. The feature search actually improved rt with more distractors.
To make a grouped bar graph we will enter the following code:
```{r}
f<-ggplot(temp, aes(x=distractors,
                    y=means,
                    fill=type))+
  geom_bar(stat="identity",position=position_dodge())+
  scale_fill_manual(values=c("darkgreen","red"),
                     name="type",
                     breaks=c("conjunction","feature"),
                     labels=c("conjunction", "feature"))+
  geom_errorbar(aes(ymax=means+sems,
                    ymin=means-sems),
                width=.2,
                position=position_dodge(.9))+
  ggtitle("Effect of Number of Distractors on Response Time")+
  labs(x="Number of Distractors",y="Response Time")+
  theme(plot.title=element_text(size=15,face="bold",vjust=.5))+
  theme(axis.title.x=element_text(size=12,face="bold",vjust=-.25))+
  theme(axis.title.y=element_text(size=12,face="bold",vjust=1))+
  theme(axis.text.x=element_text(size=10,face="bold",color="black"))+
  theme(axis.text.y=element_text(size=10,face="bold",color="black"))+
  coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),
                         max(temp$means)+2*max(temp$sems)))+
  theme(panel.border=element_blank(),axis.line=element_line())+
  theme(panel.grid.major.x=element_blank())+
  theme(panel.grid.major.y=element_line(color="darkgrey"))+
  theme(panel.grid.minor.y=element_blank())+
  theme(legend.position=c(.4,.76))+
  theme(legend.background=element_blank())+
  theme(legend.background=element_rect(color="black"))+
  theme(legend.title=element_blank())+
  theme(legend.title=element_text(size=12))+
  theme(legend.title.align=.5)+
  theme(legend.text=element_text(size=10,face="bold"))
f
```
This graph really illustrates the differences in response times, giving a clear picture of the effects of distractors on response time in both types of searches.Conjunction searches are effected much more by distractors than feature searches.

<!--
Interpret the statistical results in light of the figures.
-->


<!--
Perform a within-subjects two-way ANOVA examining for the effects
of and interactions between number of distractors and search type.
Tip: use subject as the subject's variable for this analysis.
Interpret the results of the analysis based on the statistical
results and your earlier figures.
-->
Next we will perform a within-subjects two way ANOVA examining for the effects and interactions between number of distractors and search type.
```{r}
summary(aov(rt~distractors*type+
              Error(subject/(distractors*type)),
                    data=dat))
```


<!--
Finally conduct and interpret the results of a mixed ANOVA in which
number of distractors varies within subjects and the type of search
varies between subjects. Make sure to interpret your results. Tip:
use subject2 as your subject's variable for this analysis.
-->
Finally, we will conduct a mixed ANOVA in which number of distractors varies within subjects and the type of search varies between subjects.
```{r}
summary(aov(rt~distractors*type+
              Error(subject2/distractors),
            data=dat))
```
