##Example 4

<!--
Set the working directory.
No Rchunk is needed to accomplish this.
-->

<!--
Read in the example4.csv data. For each of the following steps you
should preface your R commands with some plain English explanation.
-->

The first step is to read in the data.
```{r
dat<-read.csv("example4.csv")
```
<!--
Check the structure of the data.
-->

The next step is to check out the structure

```{r}
str(dat)
```

As we can see dat is a data frame with two variables: gender and major.

<!--
Create a table that shows how many people fell into each category.
-->

Now I'll create a table that shows how many people fell into each
category.

```{r}
table(dat$major, dat$gender)
```

From examining the table, it appears that there might more female
psych students and more male engineering statements.

NOTE: after I created the table, I ran the code to be able to
interpret the output. I also do this for the statistical analyses.
I recommend that you do this as you won't be able to draw the
appropriate conclusions or report the right data unless you go
back and forth running code and then interpreting the results.

<!--
Perform a chi-square test to determine whether or not gender and
major are independent or related. Also, interpret the results of
your test and provide a a statistical statement that supports your
interpretation. As there's no way to insert the chi-square symbol,
you can just use chi-square instead (see below).
-->

To determine if the data support the appearance of a relationship
between gender and major, I performed a chi-square test.

```{r}
chisq.test(dat$major,dat$gender)
```

According to the results of the chi-square test, major and gender
are independent of one another Chi-square(1) = 2.4, p = .12.
