## Homework 4

<!--
Remember you can use the example4.Rmd or other files to find examples
of the codes you'll need to use to complete this assignment. I strongly
urge you to copy and paste relevant code that you know works from my
examples or instructions and then edit that code so you're working
with the appropriate variables for the homework.
-->

<!--
Set the working directory.
No Rchunk is needed to accomplish this.
-->
 setwd("C:/Users/tamara/Desktop/Capstone Reproducibility/Git project")
<!--
Read in the homework4.csv data. For each of the following steps you
should preface your R commands with some plain English explanation.
-->
```{r}
 dat<-read.csv("homework4.csv")
```
This tells R to read the homework file
<!--
Check the structure of the data.
-->
```{r}
 str(dat)
 ```
This shows us that there is 80 obs with two variables:
condition factor with two levelss: control and experimental
agreement factor with two levels: yes or no


Create a table that shows how many people fell into each category.

In order to create a table you must set seed and assign the variables numeric value
```{r}
 set.seed(124)
```
Then you create a table with the following command:
```{r}
 mytable<-table(dat$gender,dat$major)
 ```
 This creates the table, but to read it you need to enter:
 ```{r}
 mytable
 ```
This shows the table of our data:

               no yes
  control      20  20
  experimental 14  26
>

<!--
Perform a chi-square test to determine whether or not the experimental
conditions and agreement with a statement are independent or related.
Also, interpret the results of your test and provide a a statistical
statement that supports your interpretation. As there's no way to
insert the chi-square symbol, you can just use chi-square instead
(see example4.Rmd).
-->
```{r}
 chisq.test(dat$condition,dat$agreement)
```
	Pearson's Chi-squared test with Yates' continuity
	correction

data:  dat$condition and dat$agreement
X-squared = 1.2788, df = 1, p-value = 0.2581

According to the data provided we would reject the null hypothesis that the data are related. There is not significant data to support that the experimental conditions are related to agreement with a statement.
p = .26 and chisq = 1.28

