## Example 5

<!--
Set the working directory.
No Rchunk is needed to accomplish this.
-->

<!--
Read in the example5.csv data. For each of the following steps you
should preface your R commands with some plain English explanation.
-->

The first step is to read in the data.
```{r}
dat<-read.csv("example5.csv")
```
<!--
Check the structure of the data.
-->

The next step is to check out the structure

```{r}
str(dat)
```

As we can see dat is a data frame with two variables: group and rt.

<!--
Find the mean response times for the experimental and control groups
-->

```{r}
mean(dat$rt[dat$group=="experimental"])
mean(dat$rt[dat$group=="control"])
```

NOTE: after I enter bits of code, I run the code or knit the document
to see what the results were of my analyses. Thus, I'm going back and
forth between running code and checking the output to make sure it
appears correct.

By looking at the means it appears that the experimental group
completed the task faster than the control group.

<!--
Run an independent t-test to determine if the mean rt for the
experimental group is different from the mean rt from the control
group. Interpret the results of your analysis after you run it.
-->

Now I'll perform an independent t-test to see if my intuition was
correct.

```{r}
t.test(dat$rt[dat$group=="control"],
       dat$rt[dat$group=="experimental"],
       var.equal=TRUE)
```

Examining the results of our t-test we can see that individuals in
the experimental group performed the task quicker than individuals
in the control group, t(48) = 3.15, p = .0028.

<!--
Run another independent t-test with the alternative hypothesis that
the experimental group is faster than the control group. Interpret
the results of your analyses.
-->

Now, I'll run an independent t-test with the alternative hypothesis
that the experimental group is faster than the control group.

```{r}
t.test(dat$rt[dat$group=="control"],
       dat$rt[dat$group=="experimental"],
       var.equal=TRUE,
       alternative="greater")
```

Examining the results of our t-test we can see that individuals in
the experimental group performed the task quicker than individuals
in the control group, t(48) = 3.15, p = .0014.
<!--
Create a bar graph of the data.
-->

In order to create a figure of the data, I need to load the
ggplot2, gplots, and dplyr libraries.


```{r}
library(ggplot2)
library(gplots)
library(dplyr)
```

Next I need to use dplyr to summarize the data and find the means and
standard errors of the means for the different groups.

```{r}
temp<-dat%>%group_by(group)%>%
  summarize(means=mean(rt), sems=sd(rt)/sqrt(length(rt)))
```

Armed with the summarized data I can now create a figure of the data.

```{r}
f<-ggplot(temp,aes(x=factor(group),y=means))+
  geom_bar(stat="identity", color="black",
           fill=c("deepskyblue2", "deeppink"))+
  geom_errorbar(aes(ymax=means+sems,
                    ymin=means-sems),
                width=.2)+
  ggtitle("Response Times For the Groups")+
  labs(x="Group", y="Response Time (ms)")+
  scale_x_discrete(breaks=c("control","experimental"),
                   labels=c("Control","Experimental"))+
  theme(plot.title=element_text(size=15,
                                face="bold",
                                vjust=.5))+
  theme(axis.title.x=element_text(size=12,
                                  face="bold",
                                  vjust=-.25))+
  theme(axis.title.y=element_text(size=12,
                                  face="bold",
                                  vjust=.25))+
  theme(axis.text.x=element_text(size=10,
                                 face="bold",
                                 color="black"))+
  theme(axis.text.y=element_text(size=10,
                                 face="bold",
                                 color="black"))+
  coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),
                         max(temp$means)+2*max(temp$sems)))+
  theme(panel.border=element_blank(),
        axis.line=element_line())+
  theme(panel.grid.major.x=element_blank())+
  theme(panel.grid.major.y=element_line(color="darkgrey"))+
  theme(panel.grid.minor.y=element_blank())
f
```

<!--
Conduct and interpret a two-tailed paired t-test on the data.
-->

```{r}
t.test(dat$rt[dat$group=="control"],
       dat$rt[dat$group=="experimental"],
       var.equal=TRUE,
       paired=TRUE)
```

Individuals in the experimental condition completed the task more
quickly than individuals in the control condition, t(24) = 8.01,
p < .05.
